---
title: "Compare new and old 2021 data Public Data runs"
output: pdf_document
author: Emily Markowitz (emily.markowitz@noaa.gov)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, comment=FALSE)
```


```{r load_functions, echo = FALSE}
# Load libraries, establish functions, and load external data
PKG <- c(
  # Keeping Organized
  "here",
  # "arsenal",
  
  # plotting
  "ggplot2",
  "rnaturalearth",
  "rnaturalearthdata",

  # other tidyverse
  "tidyr",
  "dplyr",
  "magrittr",
  "readr",
  
  # Text Management
  "stringr")


PKG <- unique(PKG)
for (p in PKG) {
  if(!require(p,character.only = TRUE)) {
    install.packages(p)
    require(p,character.only = TRUE)}
}


# Where/What are differences between columns
aa <- function(dat, colx, coly) {
  a <- dat[which(!(dat[,colx] == dat[,coly])),c(ifelse(colx != "common_name.x", "common_name.x", "scientific_name.x"), "species_code", "year", "srvy", colx, coly)] #%>% 
    # dplyr::rename(paste0(colx, ' | ', coly) = compare)
 if (nrow(a)>0) {
  if (is.character(unlist(dat[,colx])[1][[1]])) {
    a <- a %>%
      dplyr::mutate(compare = paste0(get(colx), " | ", get(coly)))
    out <- table(a$compare)
  } else if (is.numeric(unlist(dat[,colx])[1][[1]])) {
    out <- a %>%
      dplyr::mutate(difference = get(colx) - get(coly)) # %>% 
      # dplyr::select(species_code, difference)
    if (nrow(out)>100) {
      out$srvy <- as.factor(out$srvy)
      out$common_name.x <- as.factor(out$common_name.x)
      out$species_code <- as.factor(out$species_code)
      out <- summary(out)
    }
  }
 } else {
   out <- ""#paste0("No differences between ", colx, " and ", coly, ".")
 }
  return(out)
}

# find how prevelant issues in data are
wheres_trouble <- function(dat, col, trouble) {
  
  temp <- dat %>% 
    dplyr::rename(col = all_of(col)) #%>% 
    # dplyr::select(col)
  
  if (is.na(trouble)) {
  temp <- temp %>% 
    dplyr::filter(is.na(col))
  } else {
      temp <- temp %>% 
    dplyr::filter(col == trouble)
  }
  
  if (nrow(temp) == 0) {
    out <- ""#paste0("No '", trouble, "' were found in the column ", col, ".")
  } else {
    temp1 <- temp
    out <- list(#condition = paste0("Searching for '", trouble, "' in the column ", col, "."), 
              how_many = table(temp[,c("year", "srvy")])#, 
              # data_subset = temp1
              )
    names(out) <- paste0("Yes, '", trouble, "' occured ",nrow(temp)," times in the column '", col, "'.")
    
  }
  return(out)
}


plot_stations <- function(dat, srvy0) {
  world <- map_data("world2")

  dat0 <- dat %>% 
               dplyr::filter(srvy == srvy0) %>% 
               dplyr::mutate(longitude_dd = 
                               ifelse(longitude_dd>0, 
                                      abs(longitude_dd-360), 
                                      abs(longitude_dd+360))) %>%
               dplyr::select(year, longitude_dd, latitude_dd) %>% 
               unique()
  
gg <- ggplot(data = dat0,
             mapping = aes(x = longitude_dd, y = latitude_dd, group = factor(year))) +
  coord_fixed(1.3) + 
  geom_polygon(data = world,
               mapping=aes(x=long, y=lat, group=group), 
               color=NA, fill="lightgreen") +
  geom_point(size = .05) +
  ggplot2::ylim(range(dat0$latitude_dd)) + 
  ggplot2::xlim(range(dat0$longitude_dd)) 

if (length(unique(dat0$year))>20) {
gg <- gg + 
  facet_wrap("year", ncol = 4) 
} else {
# } else if (length(unique(dat0$year))>7) {
  gg <- gg +
  facet_wrap("year", ncol = 2)
# } else {
#   gg <- gg +
#   facet_wrap("year", ncol = 1)
}
gg <- gg +
  ggtitle(paste0(srvy0)) +
  theme(panel.grid.major = element_line(colour = "grey90"), 
        panel.background = element_rect(fill = NA,
                                      colour = NA),
        panel.border = element_rect(fill = NA,
                                  colour = "grey20"),
        strip.background = element_blank(), 
        strip.text = element_text(size = 5, face = "bold"),
        axis.text = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank())

return(gg)
}

```

# 2021 old public data: Load and compile data

column names have been changed to match with the new data column names

```{r load_old, echo = FALSE}

if (FALSE) {
# This year's data: 
a<-list.files(path = here::here("data", "publicdata", "2021"))
df.ls <- list()
for (i in 1:length(a)){
  b <- read_csv(file = paste0(here::here("data", "publicdata", "2021", a[i])))
  b$file <- a[i]
  df.ls[[i]]<-b
  names(df.ls)[i]<-a[i]
}

# 2021 old public data: Basic data clean

data <- SameColNames(df.ls) %>% 
  dplyr::mutate(srvy = toupper(str_extract(string = file, pattern = "[a-zA-Z]+"))) %>% 
  dplyr::filter(!is.na(year))

dataa<-data %>% 
  dplyr::select(-file)

write_csv(x = dataa, file = here::here("output", Sys.Date(),  "ak_gf_survey_public_2021.csv"))

} else {
  data <- readr::read_csv(file = paste0("../data/publicdata/ak_gf_survey_public_2021.csv"))
}

data_old <- data %>% 
  dplyr::rename(surface_temperature_c = surf_temp, 
                bottom_temperature_c = bot_temp, 
                longitude_dd = longitude, 
                latitude_dd = latitude, 
                depth_m = bot_depth, 
                date = datetime, 
                cpue_kgha = wtcpue, 
                cpue_noha = numcpue, 
                species_code = sid, 
                common_name = common, 
                scientific_name = scientific) %>%
  dplyr::mutate(survey = case_when(
    srvy == "NBS" ~ "Northern Bering Sea", 
    srvy == "EBS" ~ "Southeastern Bering Sea", 
    srvy == "BSSLOPE" ~ "Bering Sea Slope", 
    srvy == "GOA" ~ "Gulf of Alaska", 
    srvy == "AI" ~ "Aleutian Islands")) %>% 
  dplyr::mutate(
    date = format(as.Date(date, format = "%m/%d/%Y"), "%Y-%m-%d"), 
    common_name = gsub(pattern = "  ", replacement = " ", 
                       x = trimws(common_name), fixed = TRUE), 
    scientific_name = gsub(pattern = "  ", replacement = " ", 
                           x = trimws(scientific_name), fixed = TRUE))

data_old %>% 
  dplyr::mutate(srvy = as.factor(srvy), 
                survey = as.factor(survey), 
                common_name = as.factor(common_name), 
                scientific_name = as.factor(scientific_name) ,
                station = as.factor(station), 
                # vessel_name = as.factor(vessel_name), 
                date = as.factor(date)) %>%
  summary()
```

# 2021 new public data: Load and compile data

```{r load_new, echo = FALSE}
if (FALSE){
  data_new <- readr::read_csv(file = paste0(".", dir_out, "cpue_biomass_station.csv"))
}
data_new %>% 
  dplyr::mutate(srvy = as.factor(srvy), 
                survey = as.factor(survey), 
                common_name = as.factor(common_name), 
                scientific_name = as.factor(scientific_name) ,
                station = as.factor(station), 
                # vessel_name = as.factor(vessel_name), 
                date = as.factor(date)) %>%
  summary()
```

# Bind new and old 2021 data to see where differences are

 - -9999 --> NA from cpue_noha and cpue_kgha in the old data so we can compare new and old more easily
 - commas (",") removed from old data (e.g., commas in "Clark, 2006" vs "Clark 2006") because I bet it doesn't matter. **Should the comma stay for the final product?**

```{r bind_new_old}
dat <- dplyr::full_join(
  x = data_new %>% 
    dplyr::mutate(dataset = "new", 
                  scientific_name = gsub(pattern = ",", replacement = "", 
                                         x = scientific_name, fixed = TRUE)), 
  y = data_old %>% 
    dplyr::mutate(dataset = "old", 
                  cpue_kgha = ifelse(cpue_kgha == -9999, NA, cpue_kgha), 
                  cpue_noha = ifelse(cpue_noha == -9999, NA, cpue_noha)),
  by = c("year", "vessel_id", "stratum", "station", "species_code", "haul", "cruise",  "srvy"))
dat
```

## What observations are in new and not in old, visa versa

column dataset = 
 - both = observations are in both new and old datasets
 - old = observations are only in the old dataset
 - new = observations are only in the new dataset

```{r id_dataset_origin}
dat0 <- dat %>% 
  dplyr::mutate(dataset = dplyr::case_when(
    (dataset.x == "new" & dataset.y == "old") ~ "both", 
    dataset.y == "old" ~ "old", 
    dataset.x == "new" ~ "new", 
    TRUE ~ "neither"))
dat0 %>% 
  dplyr::select(dataset) %>% 
  table()
```

## datasets only in the new data

### year by survey

```{r new_yr_srvy}
dat0 %>% 
  dplyr::filter(dataset == "new") %>% # data only in new data, not in old data
  dplyr::select(year, srvy) %>% 
  table()
```

### species_code by survey

```{r new_spp_srvy}
dat0 %>% 
  dplyr::filter(dataset == "new") %>% # data only in new data, not in old data
  dplyr::select(species_code, srvy) %>% 
  table()
```

## datasets only in the old data

### year by survey

```{r old_yr_srvy}
dat0 %>% 
  dplyr::filter(dataset == "old") %>% # data only in old data, not in new data
  dplyr::select(year, srvy) %>% 
  table() 
```

### species_code by survey

```{r old_spp_srvy}
dat0 %>% 
  dplyr::filter(dataset == "old") %>% # data only in old data, not in new data
  dplyr::select(species_code, srvy) %>% 
  table() 
```

## Look through new and old 2021 data and note where differences are

Please let me know if there are any differences between these datasets that are unacceptable (e.g., should be more like the old data or completely changed) and what tables and columns I should reference to fix these issues. 

Note that this will only print where these trouble makers occurred if they occurred. 

 - characters: x = new; y = old; new | old
 - numbers: a summary table of where these differences occurred with additional columns for context (common_name, species_code, year, and srvy)


**Things to look out for in scientific names** (that we might want to make formative changes in "RACEBASE.SPECIES_CLASSIFICATION$report_name_scientific"): 

 - NOTE I already removed commas (see note above; bet it doesn't matter) from old data
 - "unid." and "uid" in new data vs [nothing] in old data
 - where "egg"/"egg case" is included in scientific names of old data and if it should be included in the new data. Should these "egg"/"egg case" notes only be included in the common name? 
 - ownership dispute between McLean and Clark (e.g., "Neptunea sp. E (McLean and Clark)" and "Neptunea sp. E (Clark and McLean)" and ..."(McLean & Clark))


**Other notes**

 - -9999 --> NA from cpue_noha and cpue_kgha in the old data so we can compare new and old more easily

```{r differences_new_old}
col <- gsub(pattern = ".x", replacement = "", 
            x = names(dat)[grep(pattern = ".x", x = names(dat), fixed = TRUE)])
col <- col[-length(col)] 
# # Compare x to y
for(i in 1:length(col)) {
  a <- aa(dat = dat, colx = paste0(col[i], ".x"), coly = paste0(col[i], ".y"))
  if (class(a) == "table") {
     print(paste0("------------- Differences between '", col[i], ".x' and '", col[i], ".y' -------------"))
     print(aa(dat = dat, colx = paste0(col[i], ".x"), coly = paste0(col[i], ".y")))  
  }
}
```

# What did the new 2021 data look like?

## Look through new 2021 data and see if there were any -9999, NA, "", or 0's

Note that this will only print where these trouble makers occurred if they occurred. 

```{r trouble_makers}
trouble <- c(-9999, NA, "", 0)
cols <- names(data_new) 
cols <- cols[cols != "date"]

for (i in 1:length(trouble)){
  print(paste0("------------- Are there any '", trouble[i], "' in these columns? -------------"))
  for (ii in 1:length(cols)) {
     a <- wheres_trouble(dat = data_new, col = cols[ii], trouble = trouble[i])
      if (class(a) == "list") {
        print(a) 
      }
  }
}
```

## What stations for each year and survey look like

```{r station_map, fig.height = 10, fig.width = 7}
for (i in 1:length(unique(data_new$srvy))) {
  print(plot_stations(dat = data_new, srvy0 = unique(data_new$srvy)[i]))
}
```


