---
title: "Check Public Data Output"
output: 
  word_document:
    reference_docx: word_styles.docx
author: Emily Markowitz (emily.markowitz@noaa.gov)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  message = FALSE, 
  comment = FALSE)
```


```{r load_functions, echo = FALSE}
source("../code/check_functions.R")
```

# 2021 OLD public data: Load and compile data

column names have been changed to match with the new data column names

**Note:** 

-9999s -> NA in surface_temperature_c

-9999s -> NA in bottom_temperature_c

-9999s -> NA in cpue_kgha

-9999s removed from cpue_noha

```{r load_old, echo = FALSE}

if (FALSE) {
  
  # This year's data: 
  a<-list.files(path = here::here("data", "publicdata", "2021"))
  df.ls <- list()
  for (i in 1:length(a)){
    b <- read_csv(file = paste0(here::here("data", "publicdata", "2021", a[i])))
    b$file <- a[i]
    df.ls[[i]]<-b
    names(df.ls)[i]<-a[i]
  }
  
  # 2021 old public data: Basic data clean
  public_old <- SameColNames(df.ls) %>% 
    dplyr::mutate(srvy = toupper(str_extract(string = file, pattern = "[a-zA-Z]+"))) %>% 
    dplyr::filter(!is.na(year)) 
  
  write_csv(x = public_old, file = here::here("data", "publicdata", "ak_gf_survey_public_2021.csv"))
  
} else {
  public_old <- readr::read_csv(file = here::here("data", "publicdata", "ak_gf_survey_public_2021.csv"))
}

public_old <- public_old0 <- public_old %>%
  dplyr::rename(surface_temperature_c = surf_temp, 
                bottom_temperature_c = bot_temp, 
                vessel_id = vessel_id, 
                longitude_dd_start = longitude, 
                latitude_dd_start = latitude, 
                depth_m = bot_depth, 
                date_time = datetime, 
                cpue_kgha = wtcpue, 
                cpue_noha = numcpue, 
                species_code = sid, 
                common_name = common, 
                scientific_name = scientific) %>%
  dplyr::mutate(srvy = ifelse(srvy == "BSSLOPE", "BSS", srvy)) %>%
  dplyr::mutate(survey = case_when(
    srvy == "NBS" ~ "Northern Bering Sea", 
    srvy == "EBS" ~ "Southeastern Bering Sea", 
    srvy == "BSS" ~ "Bering Sea Slope", 
    srvy == "GOA" ~ "Gulf of Alaska", 
    srvy == "AI" ~ "Aleutian Islands")) %>% 
  dplyr::filter(# to match the public_new
    year != 2020 & # no surveys happened this year because of COVID
      (year >= 1982 & srvy %in% c("EBS", "NBS") | # 1982 BS inclusive - much more standardized after this year
         srvy %in% "BSS" | # keep all years of the BSS
         year >= 1991 & srvy %in% c("AI", "GOA")) ) %>%  # 1991 AI and GOA (1993) inclusive - much more standardized after this year
  dplyr::mutate(date_time = 
                  as.character(format(as.Date(date_time, format="%m/%d/%Y"),"%m/%d/%Y")), 
                common_name = gsub(pattern = "  ", replacement = " ", 
                                   x = trimws(common_name), fixed = TRUE), 
                scientific_name = gsub(pattern = "  ", replacement = " ", 
                                       x = trimws(scientific_name), fixed = TRUE), 
                surface_temperature_c = ifelse(surface_temperature_c == -9999, NA, surface_temperature_c), 
                bottom_temperature_c = ifelse(bottom_temperature_c == -9999, NA, bottom_temperature_c), 
                cpue_kgha = ifelse(cpue_kgha == -9999, NA, cpue_kgha), 
                cpue_noha = ifelse(cpue_noha == -9999, NA, cpue_noha),
                # count = NA, 
                # weight_kg = NA,
                area_swept_ha = NA) 

public_old %>% 
  dplyr::mutate(srvy = as.factor(srvy), 
                survey = as.factor(survey), 
                common_name = as.factor(common_name), 
                scientific_name = as.factor(scientific_name) ,
                station = as.factor(station), 
                # vessel_name = as.factor(vessel_name), 
                date_time = as.factor(date_time)) %>%
  summary()

readr::write_csv(x = public_old, file = paste0(".", dir_out, "/check/public_old.csv"))

```

## Check public_old for overarching structure and summarization issues

```{r public_old_check}

# data_check(dat = public_old0) # run checks to see how the data looks, but don't implement it

public_old <- public_old %>%
  dplyr::filter(
    cpue_kgha != 0 & # removed because doesnt exist in the public_new
      stratum != 0 & # removed because doesnt exist in the public_new
      !is.na(species_code)) # removed because doesnt exist in the public_new

public_old %>% 
  dplyr::mutate(srvy = as.factor(srvy), 
                survey = as.factor(survey), 
                common_name = as.factor(common_name), 
                scientific_name = as.factor(scientific_name) ,
                station = as.factor(station), 
                # vessel_name = as.factor(vessel_name), 
                date_time = as.factor(date_time)) %>%
  summary()
```


Cleaned data! How many rows were removed using the data_check()

```{r public_old_check_summary, echo = FALSE}
# print(paste0("how many rows were removed using the data_check()?: ", nrow(public_old0)-nrow(public_old) ))
# public_old %>% 
#   dplyr::mutate(srvy = as.factor(srvy), 
#                 survey = as.factor(survey), 
#                 common_name = as.factor(common_name), 
#                 scientific_name = as.factor(scientific_name) ,
#                 station = as.factor(station), 
#                 # vessel_name = as.factor(vessel_name), 
#                 date_time = as.factor(date_time)) %>%
#   summary()
```


# 2021 NEW public data: Load and compile data

```{r load_new, echo = FALSE}

public_new <- readr::read_csv(file = paste0(".", dir_out, "cpue_station.csv"))

public_new <- public_new %>% 
  dplyr::mutate(date_time = as.character(format(as.Date(date_time,
                                                        format="%m/%d/%Y"),"%m/%d/%Y")))

public_new %>% 
  dplyr::mutate(srvy = as.factor(srvy), 
                survey = as.factor(survey), 
                common_name = as.factor(common_name), 
                scientific_name = as.factor(scientific_name) ,
                station = as.factor(station), 
                date_time = as.factor(date_time)) %>% 
  summary()
```

## Check public_new for overarching structure and summarization issues

```{r new_public_check}
public_new0<-public_new # the OG data

# data_check(public_new0) # run checks to see how the data looks, but don't implement it
```

Cleaned data! How many rows were removed using the data_check()

```{r new_public_check_summary, echo = FALSE}
# print(paste0("how many rows were removed using the data_check()?: ", nrow(public_new0)-nrow(public_new) ))
# public_new %>% 
#   dplyr::mutate(srvy = as.factor(srvy), 
#                 survey = as.factor(survey), 
#                 common_name = as.factor(common_name), 
#                 scientific_name = as.factor(scientific_name) ,
#                 station = as.factor(station), 
#                 # vessel_name = as.factor(vessel_name), 
#                 date_time = as.factor(date_time)) %>%
#   summary()
```


## What does the new 2021 data look like?

Look through new 2021 data and see if there were any -9999, NA, "", or 0's

Note that this will only print where these trouble makers occurred if they occurred. 

```{r trouble_makers}
trouble <- c(-9999, NA, "", 0)
cols <- c("stratum", "station", "species_code", 
          "cpue_kgha", "cpue_noha", "weight_kg", "count", 
          "bottom_temperature_c", "surface_temperature_c", "depth_m")

for (i in 1:length(trouble)){
  print(paste0("------------- Are there any '", trouble[i], "' in these columns? -------------"))
  for (ii in 1:length(cols)) {
    a <- wheres_trouble(dat = public_new, col = cols[ii], trouble = trouble[i])
    if (class(a) == "list") {
      print(a) 
    }
  }
}
```

# Bind new and old 2021 data to see where differences are

- -9999 --> NA from cpue_noha and cpue_kgha in the old data so we can compare new and old more easily
- commas (",") removed from old data (e.g., commas in "Clark, 2006" vs "Clark 2006") because I bet it doesn't matter. **Should the comma stay for the final product?**

```{r bind_new_old}
dat <- dplyr::full_join(
  x = public_new %>% 
    dplyr::mutate(dataset = "new", 
                  scientific_name = gsub(pattern = ",", replacement = "", 
                                         x = scientific_name, fixed = TRUE)), 
  y = public_old %>% 
    dplyr::mutate(dataset = "old", 
                  count = NA, 
                  weight_kg = NA),
  by = c("year", "vessel_id", "stratum", "station", "species_code", "haul", "cruise",  "srvy")) %>% 
  dplyr::mutate(dataset = dplyr::case_when(
    (dataset.x == "new" & dataset.y == "old") ~ "both", 
    dataset.y == "old" ~ "old", 
    dataset.x == "new" ~ "new", 
    TRUE ~ "neither")) %>% 
  dplyr::select(-dataset.y, -dataset.x, -starts_with("survey."))

dat
```

## What observations are in new and not in old, visa versa

column dataset = 
- both = observations are in both new and old datasets
- old = observations are only in the old dataset
- new = observations are only in the new dataset

```{r id_dataset_origin}
dat0 <- dat
dat0 %>% 
  dplyr::select(dataset) %>% 
  table()
```

## datasets only in the new data

### year by survey

```{r new_yr_srvy}
dat0 %>% 
  dplyr::filter(dataset == "new") %>% # data only in new data, not in old data
  dplyr::select(year, srvy) %>% 
  table()
```

### species_code by survey

```{r new_spp_srvy}
# dat0 %>% 
#   dplyr::filter(dataset == "new") %>% # data only in new data, not in old data
#   dplyr::select(species_code, srvy) %>% 
#   table()
```

## datasets only in the old data

### year by survey

```{r old_yr_srvy}
dat0 %>% 
  dplyr::filter(dataset == "old") %>% # data only in old data, not in new data
  dplyr::select(year, srvy) %>% 
  table() 
```

### species_code by survey

```{r old_spp_srvy}
# dat0 %>% 
#   dplyr::filter(dataset == "old") %>% # data only in old data, not in new data
#   dplyr::select(species_code, srvy) %>% 
#   table() 
```


## CPUE for major species

```{r CPUE_kgha_plot_new_old, fig.height = 7, fig.width = 4.8}
spp <- c(69323, # blue king crab 
         68580, # snow crab
         69322, # red king crab
         10110, # arrowtooth flounder
         21720, # Pacific cod
         21740, # walleye pollock
         10112, # Kamchatka flounder
         471, # Alaska skate
         10285, # Alaska plaice
         10210, # yellowfin sole
         10261 # northern rock sole
)

# for (i in 1:length(unique(dat$srvy))) {
for (ii in 1:length(spp)){
  print(plot_cpue(dat = dat, spp0 = spp[ii], 
                      dataset1 = "public_new", 
                      dataset2 = "public_old"))
}

```

## Look through new and old 2021 data and note where differences are

Please let me know if there are any differences between these datasets that are unacceptable (e.g., should be more like the old data or completely changed) and what tables and columns I should reference to fix these issues. 

Note that this will only print where these trouble makers occurred if they occurred. 

- characters: x = new; y = old; new | old
- numbers: a summary table of where these differences occurred with additional columns for context (common_name, species_code, year, and srvy)


**Things to look out for in scientific names** (that we might want to make formative changes in "RACEBASE.SPECIES_CLASSIFICATION$report_name_scientific"): 

- NOTE I already removed commas (see note above; bet it doesn't matter) from old data
- "unid." and "uid" in new data vs [nothing] in old data
- where "egg"/"egg case" is included in scientific names of old data and if it should be included in the new data. Should these "egg"/"egg case" notes only be included in the common name? 
- ownership dispute between McLean and Clark (e.g., "Neptunea sp. E (McLean and Clark)" and "Neptunea sp. E (Clark and McLean)" and ..."(McLean & Clark))


**Other notes**

- -9999 --> NA from cpue_noha and cpue_kgha in the old data so we can compare new and old more easily

```{r differences_new_old}
col <- gsub(pattern = ".x", replacement = "", 
            x = names(dat)[grep(pattern = ".x", x = names(dat), fixed = TRUE)])
col <- col[-length(col)] 
# # Compare x to y
for(i in 1:length(col)) {
  a <- diff_between_cols(dat = dat, colx = paste0(col[i], ".x"), coly = paste0(col[i], ".y"))
  if (class(a) == "table") {
    print(paste0("----- Differences between '", col[i], ".x' (new) and '", col[i], ".y' (old) -----"))
    print(diff_between_cols(dat = dat, colx = paste0(col[i], ".x"), coly = paste0(col[i], ".y")))  
  }
}
```


```{r diff_xnew_yold}
dat0 <- dat %>% 
  dplyr::mutate(
    dplyr::across(dplyr::starts_with("cpue_"), round, digits = 2), 
    cpue_kgha_diff = cpue_kgha.x-cpue_kgha.y, 
    cpue_noha_diff = cpue_noha.x-cpue_noha.y) %>% 
  dplyr::select(year, srvy, cruise, stratum, station, haul, vessel_id, 
                species_code, common_name.x, scientific_name.x, 
                cpue_kgha.x, cpue_kgha.y, cpue_kgha_diff, 
                cpue_noha.x, cpue_noha.y, cpue_noha_diff) %>% 
  dplyr::filter(cpue_kgha_diff >= 0.01 | cpue_noha_diff >= 0.01) %>% 
  dplyr::arrange(-cpue_kgha_diff, -cpue_noha_diff)

readr::write_csv(dat0, 
                 file = paste0(".", dir_out, "/check/diff_xnew_yold.csv"))

summary(dat0)
```

# Compare with our design-based CPUE estimates?

Here I am compiling the EBS, NBS, GOA, AI verified, design-based datasets

```{r format_design_based}

design_based1 <- dplyr::bind_rows(
  readr::read_csv(file = "../data/oracle/cpue_ai.csv"), 
  readr::read_csv(file = "../data/oracle/cpue_goa.csv")) %>% 
  janitor::clean_names() %>% 
  dplyr::left_join(
    x = ., 
    y = haul0 %>% 
      dplyr::select(hauljoin, stationid, start_latitude, start_longitude), 
    by = "hauljoin")  %>% 
  dplyr::left_join(
    x = ., 
    y = spp_info %>% 
      dplyr::select(common_name, scientific_name, species_code), 
    by = "species_code") %>% 
  dplyr::rename(latitude = start_latitude, 
                longitude = start_longitude, 
                weight_kg = weight, 
                count = number_fish) %>% 
  dplyr::mutate(cpue_kgha = wgtcpue/100, # denominator orig in km2
                cpue_noha = numcpue/100, # denominator orig in km2
                area_swept_ha = effort*100) %>% # denominator orig in km2
  dplyr::select(-x1, -catchjoin, -wgtcpue, -numcpue, -effort)

design_based2 <- dplyr::bind_rows(
  readr::read_csv(file = "../data/oracle/cpue_ebs_plusnw.csv") %>% 
    dplyr::mutate(survey = "EBS"), 
  readr::read_csv(file = "../data/oracle/cpue_nbs.csv") %>% 
    dplyr::mutate(survey = "NBS")) %>% 
  janitor::clean_names() %>% 
  dplyr::select(-x1)  %>% 
  dplyr::left_join(
    x = ., 
    y = haul %>% 
      dplyr::select(hauljoin, cruise, distance_fished, net_width), 
    by = "hauljoin") %>% 
  dplyr::mutate(weight_kg = NA, 
                count = NA) %>% 
  dplyr::rename(area_swept_ha = area_fished_ha, 
                scientific_name = species_name)

design_based <- dplyr::bind_rows(design_based1, design_based2) %>% 
  dplyr::rename(srvy = survey, 
                vessel_id = vessel, 
                distance_fished_km = distance_fished, 
                net_width_m = net_width, 
                station = stationid, 
                latitude_dd_start = latitude, 
                longitude_dd_start = longitude)

design_based0<-design_based
readr::write_csv(x = design_based, file = paste0(".", dir_out, "/check/design_based.csv"))

dim(design_based) # before summarizing
design_based %>% 
  dplyr::mutate(srvy = as.factor(srvy), 
                year = as.factor(year),
                common_name = as.factor(common_name), 
                scientific_name = as.factor(scientific_name) ,
                station = as.factor(station)) %>% 
  summary()
```

## Check design_based for overarching structure and summarization issues

```{r design_based_check}
design_based0<-design_based # the OG data

# data_check(
#   dat = design_based0 %>% 
#     dplyr::select(-weight_kg, -count)) # run checks to see how the data looks, but don't implement it
```

how many rows were removed using the data_check()

```{r design_based_check_summary, echo = FALSE}
# print(paste0("how many rows were removed using the data_check()?: ", nrow(design_based0)-nrow(design_based) ))
# design_based %>% 
#   dplyr::mutate(srvy = as.factor(srvy), 
#                 # survey = as.factor(survey),
#                 common_name = as.factor(common_name), 
#                 scientific_name = as.factor(scientific_name) ,
#                 station = as.factor(station)#, 
#                 # vessel_name = as.factor(vessel_name), 
#                 # date_time = as.factor(date_time)
#   ) %>%
#   summary()
```

## Compare public_new to design_based data

And now I am joining the new data and the designed based data together to see where differences arise. 

I am rounding the CPUE and area_swept columns to 3 digits so we only see the big differences

```{r join_diff_designbased_new}

dat <- dplyr::full_join(
  x = public_new %>% 
    dplyr::select(common_name, species_code, scientific_name, 
                  cpue_kgha, cpue_noha, 
                  area_swept_ha, #distance_fished_km, net_width_m, 
                  haul, cruise, station, stratum, srvy, 
                  count, weight_kg,
                  vessel_id, latitude_dd_start, longitude_dd_start, year),
  y = design_based %>% 
    dplyr::select(common_name, species_code, scientific_name, 
                  cpue_kgha, cpue_noha, 
                  area_swept_ha, #distance_fished_km, net_width_m, 
                  haul, cruise, station, stratum, srvy, 
                  count, weight_kg,
                  vessel_id, latitude_dd_start, longitude_dd_start, year), 
  by = c("year", "haul", "cruise", "srvy", "species_code", "vessel_id", "stratum", "station")) %>%
  dplyr::mutate(dplyr::across(dplyr::starts_with("cpue_"), round, digits = 3), 
                dplyr::across(dplyr::starts_with("area_swept_"), round, digits = 3)) 

col <- gsub(pattern = ".x", replacement = "", 
            x = names(dat)[grep(pattern = ".x", x = names(dat), fixed = TRUE)])

## Compare x to y
for(i in 1:length(col)) {
  a <- diff_between_cols(dat = dat, 
                         colx = paste0(col[i], ".x"), 
                         coly = paste0(col[i], ".y"))
  if (sum(class(a) %in% "tbl_df")>0) {
    print(paste0("----- Differences between '", col[i], ".x' (new public) and '", col[i], ".y' (design) -----"))
    print(diff_between_cols(dat = dat, colx = paste0(col[i], ".x"), coly = paste0(col[i], ".y")))  
  }
}

```

```{r diff_xnew_ydesign}
dat0 <- dat %>% 
  dplyr::mutate(
    cpue_kgha_diff = cpue_kgha.x-cpue_kgha.y, 
    cpue_noha_diff = cpue_noha.x-cpue_noha.y, 
    area_swept_ha_diff = area_swept_ha.x-area_swept_ha.y) %>% 
  dplyr::select(-common_name.y, -scientific_name.y) %>% 
  dplyr::filter(cpue_kgha_diff >= 0.01 | 
                  cpue_noha_diff >= 0.01 | 
                  area_swept_ha_diff >= 0.01 ) %>% 
  dplyr::arrange(srvy, cruise, haul, vessel_id, -cpue_kgha_diff, -cpue_noha_diff, -area_swept_ha_diff) %>% 
  dplyr::relocate(year, srvy, cruise, haul, station, stratum, vessel_id, 
                  species_code, common_name.x, 
                  area_swept_ha.x, area_swept_ha.y, area_swept_ha_diff, 
                  weight_kg.x, weight_kg.y, cpue_kgha.x, cpue_kgha.y, cpue_kgha_diff, 
                  count.x, count.y, cpue_kgha.x, cpue_kgha.y, cpue_kgha_diff)

readr::write_csv(dat0, 
                 file = paste0(".", dir_out, "/check/diff_xnew_ydesign.csv"))

summary(dat0)
```

## CPUE for major species

```{r CPUE_kgha_plot_new_design}
spp <- c(69323, # blue king crab 
         68580, # snow crab
         69322, # red king crab
         10110, # arrowtooth flounder
         21720, # Pacific cod
         21740, # walleye pollock
         10112, # Kamchatka flounder
         471, # Alaska skate
         10285, # Alaska plaice
         10210, # yellowfin sole
         10261 # northern rock sole
)

# for (i in 1:length(unique(dat$srvy))) {
for (ii in 1:length(spp)){
  print(plot_cpue(dat = dat %>% 
                    dplyr::filter(srvy != "BSS"), 
                  spp0 = spp[ii], 
                      dataset1 = "public_new", 
                      dataset2 = "design_based"))
}

```

# What stations for each year and survey look like in the new public data

```{r station_map}
for (i in 1:length(unique(public_new$srvy))) {
  print(plot_stations(dat = public_new %>% 
                        dplyr::rename(longitude_dd = longitude_dd_start, 
                                      latitude_dd = latitude_dd_start), 
                      srvy0 = unique(public_new$srvy)[i]))
}
```

