---
title: "Compare 2019 and new 2021 data Public Data runs"
output: pdf_document
author: Emily Markowitz (emily.markowitz@noaa.gov)
fig_width: 6 
fig_height: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries, establish functions, and load external data

```{r, warning=FALSE, message=FALSE, comment=FALSE}
library(readr)
library(arsenal)
library(magrittr)
library(stringr)
library(dplyr)
library(here)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)

#' Bind multiple data sets by column in a list and make sure columns match up
#'
#' @param df.ls A list of datasets that need to be bound together by column.
#'
#' @return data.frame
#' @export
SameColNames<-function(df.ls) {
  #All column names
  colnames0<-c()
  for (i in 1:length(df.ls)){
    df0<-df.ls[[i]]
    df0<-janitor::clean_names(df0)
    df.ls[[i]]<-df0
    colnames0<-c(colnames0, (colnames(df0)))
  }
  colnames0<-sort(unique(colnames0), decreasing = T)
  
  #New df's
  df.ls0<-list()
  df.rbind0<-c()
  for (i in 1:length(df.ls)){
    df0<-df.ls[[i]]
    colnames.out<-colnames0[!(colnames0 == colnames(df0))]
    if (length(colnames.out) != 0) {
      for (ii in 1:length(colnames.out)){
        df0[,(ncol(df0)+1)]<-NA
        names(df0)[ncol(df0)]<-colnames.out[ii]
      }
    }
    df0<-df0[,match(table =  colnames(df0), x = colnames0)]
    df.ls0[[i]]<-df0
    names(df.ls0)[i]<-names(df.ls)[i]
    df.rbind0<-rbind.data.frame(df.rbind0, df0)
  }
  return(df.rbind0)
}

# Where/What are differnces between columns
aa <- function(dat, colx, coly) {
  
  b <- which(!(dat[,colx] == dat[,coly]))
 a <- dat[b, colx]
 
 if (length(b)>0) {
 str <- paste0("Between ", colx, " and ", coly, ", ", colx, " had these values: ", paste(unlist(a), collapse = ", "), 
               " at these rows: ", 
               paste(b, collapse = ", "), ".")
 } else {
   str <- paste0("No differences between ", colx, " and ", coly, ".")
 }
  return(str)
}

# find how prevelant issues in data are
wheres_trouble <- function(dat, col, trouble) {
  
  temp <- dat %>% 
    dplyr::rename(col = all_of(col)) %>% 
    dplyr::filter(col == trouble)
  
  if (nrow(temp) == 0) {
    out <- paste0("No '", trouble, "' were found in the column ", col, ".")
  } else {
    temp1 <- temp
    out <- list(condition = paste0("Searching for ", trouble, " in the column ", col, "."), 
              how_many = table(temp[,c("year", "srvy")]), 
              data_subset = temp1)
  }

  return(out)
}


plot_stations <- function(dat, srvy0) {
# for (i in 1:length(unique(dat$srvy))){
  # srvy0 <- unique(dat$srvy)[i]

  world <- map_data("world2")

  dat0 <- dat %>% 
               dplyr::filter(srvy == srvy0) %>% 
               dplyr::mutate(longitude = 
                               ifelse(longitude>0, 
                                      abs(longitude-360), 
                                      abs(longitude+360))) %>%
               dplyr::select(year, longitude, latitude) %>% 
               unique()
  
gg <- ggplot(data = dat0,
             mapping = aes(x = longitude, y = latitude, group = factor(year))) +
  coord_fixed(1.3) + 
  geom_polygon(data = world,
               mapping=aes(x=long, y=lat, group=group), 
               color=NA, fill="lightgreen") +
  geom_point(size = .05) +
  ggplot2::ylim(range(dat0$latitude)) + 
  ggplot2::xlim(range(dat0$longitude)) + 
  facet_wrap("year") +
  ggtitle(paste0(srvy0)) +
  # theme_bw() %>%
  theme(panel.grid.major = element_line(colour = "grey90"), 
        panel.background = element_rect(fill = NA,
                                      colour = NA),
        panel.border = element_rect(fill = NA,
                                  colour = "grey20"),
        strip.background = element_blank(), 
        strip.text = element_text(size = 5, face = "bold"),
        axis.text = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank())

return(gg)
}


# from RACE_DATA.VESSELS
vessels0 <- read_csv(file = paste0(here::here("data", "oracle", "vessels.csv")))

```

# 2019: Load and compile data (for comparison to 2021 data)

```{r, warning=FALSE, message=FALSE, comment=FALSE}
# Last year's data:
# installed from: https://apps-afsc.fisheries.noaa.gov/RACE/groundfish/survey_data/data.htm
# on 11/18/2021
# d2019 <- read_csv(file = paste0(dir_data, "/publicdata/all_public_race_gf_data.csv"))

a<-list.files(path = here::here("data", "publicdata", "GF", "2019"), pattern = ".zip")
df.ls <- list()
for (i in 1:length(a)){
  unzip(zipfile = paste0(here::here("data", "publicdata", "GF", "2019", a[i])),
        overwrite = TRUE,
        exdir = paste0(here::here("data", "publicdata", "GF", "2019",
                                  gsub(pattern = ".zip", replacement = "", a[i], fixed = TRUE))))
  file.copy(from = paste0(here::here("data", "publicdata", "GF", "2019", 
                                  gsub(pattern = ".zip", replacement = "", a[i], fixed = TRUE), 
                                   gsub(pattern = "zip", replacement = "csv", a[i]))), 
            to = paste0(here::here("data", "publicdata", "GF", "2019", 
                                  gsub(pattern = "zip", replacement = "csv", a[i]))))
  b <- read_csv(file = paste0(here::here("data", "publicdata", "GF", "2019", 
                                  gsub(pattern = "zip", replacement = "csv", a[i]))))
  b$file <- a[i]
  df.ls[[i]]<-b
  names(df.ls)[i]<-a[i]
  
}
```

# 2019: Basic data clean (for comparison to 2021 data)


```{r}
d2019 <- SameColNames(df.ls) %>% 
  dplyr::mutate(srvy = toupper(str_extract(string = file, pattern = "[a-zA-Z]+"))) %>% 
  dplyr::filter(!is.na(year)) %>%
  dplyr::arrange(vessel) %>% 
  dplyr::arrange(station) %>% 
  dplyr::arrange(wtcpue) %>% 
  dplyr::arrange(srvy) %>% 
  dplyr::arrange(sid) %>% 
  dplyr::arrange(year) %>% 
  dplyr::left_join(x = ., # add vessel names
                   y = vessels0 %>% 
                     dplyr::select(VESSEL_ID, NAME) %>% 
                     dplyr::mutate(NAME = stringr::str_to_title(NAME)), 
                   by = c("vessel" = "VESSEL_ID")) %>% 
  dplyr::rename(vessel_id = vessel) %>% 
  dplyr::rename(vessel = NAME) %>% 
  dplyr::mutate(survey = case_when(
    srvy == "NBS" ~ "Northern Bering Sea", 
    srvy == "EBS" ~ "Southeastern Bering Sea", 
    srvy == "BSSLOPE" ~ "Bering Sea Slope", 
    srvy == "GOA" ~ "Gulf of Alaska", 
    srvy == "AI" ~ "Aleutian Islands")) %>% 
  dplyr::relocate("year", "srvy", "survey", "haul", "cruise", 
                  "common", "sid", "scientific", 
                  "longitude", "latitude", "stratum", "station", 
                  "wtcpue", "numcpue", "surf_temp",  "bot_temp", "bot_depth", 
                  "vessel_id", "vessel", "datetime", "file")

d2019a<-d2019 %>% 
  dplyr::select(-file)

write_csv(x = d2019a, file = here::here("data", "publicdata", "ak_gf_survey_public_2019.csv"))
head(d2019)
```

# 2021: Load and compile data

```{r, warning=FALSE, message=FALSE, comment=FALSE}
# This year's data: 
a<-list.files(path = here::here("data", "publicdata", "GF", "2021"))
df.ls <- list()
for (i in 1:length(a)){
  b <- read_csv(file = paste0(here::here("data", "publicdata", "GF", "2021", a[i])))
  b$file <- a[i]
  df.ls[[i]]<-b
  names(df.ls)[i]<-a[i]
}
```

# 2021: Basic data clean

```{r}
d2021 <- SameColNames(df.ls) %>% 
  dplyr::mutate(srvy = toupper(str_extract(string = file, pattern = "[a-zA-Z]+"))) %>% 
  dplyr::filter(!is.na(year)) %>%
  dplyr::arrange(vessel) %>% 
  dplyr::arrange(station) %>% 
  dplyr::arrange(wtcpue) %>% 
  dplyr::arrange(srvy) %>% 
  dplyr::arrange(sid) %>% 
  dplyr::arrange(year) %>% 
  dplyr::left_join(x = ., # add vessel names
                   y = vessels0 %>% 
                     dplyr::select(VESSEL_ID, NAME) %>% 
                     dplyr::mutate(NAME = stringr::str_to_title(NAME)), 
                   by = c("vessel" = "VESSEL_ID")) %>% 
  dplyr::rename(vessel_id = vessel) %>% 
  dplyr::rename(vessel = NAME) %>% 
  dplyr::mutate(survey = case_when(
    srvy == "NBS" ~ "Northern Bering Sea", 
    srvy == "EBS" ~ "Southeastern Bering Sea", 
    srvy == "BSSLOPE" ~ "Bering Sea Slope", 
    srvy == "GOA" ~ "Gulf of Alaska", 
    srvy == "AI" ~ "Aleutian Islands"))  %>% 
  dplyr::relocate("year", "srvy", "survey", "haul", "cruise", 
                  "common", "sid", "scientific", 
                  "longitude", "latitude", "stratum", "station", 
                  "wtcpue", "numcpue", "surf_temp",  "bot_temp", "bot_depth", 
                  "vessel_id", "vessel", "datetime", "file") 

d2021a<-d2021 %>% 
  dplyr::select(-file)

write_csv(x = d2021a, file = here::here("data", "publicdata", "ak_gf_survey_public_2021.csv"))
head(d2021)

```

# Bind 2019 and 2021 data to see where differences are

```{r}
dat <- dplyr::full_join(
  x = d2019, 
  y = d2021[d2021$year != 2021, ], 
  by = c("year", "vessel", "stratum", "station", "sid", "haul", "cruise",  "srvy"))
head(dat)
```

# Look through 2019 and 2021 data and note where differences are

## Compare 2021 (without 2021 observations in 2021 data) to 2019

```{r}
# Compare x to y
aa(dat = dat, colx = "surf_temp.y", coly = "surf_temp.x")
aa(dat = dat, colx = "wtcpue.y", coly = "wtcpue.x")
aa(dat = dat, colx = "scientific.y", coly = "scientific.x")
aa(dat = dat, colx = "numcpue.y", coly = "numcpue.x")
aa(dat = dat, colx = "longitude.y", coly = "longitude.x")
aa(dat = dat, colx = "latitude.y", coly = "latitude.x")
aa(dat = dat, colx = "datetime.y", coly = "datetime.x")
# aa(dat = dat, colx = "file.y", coly = "file.x")
aa(dat = dat, colx = "bot_temp.y", coly = "bot_temp.x")
aa(dat = dat, colx = "bot_depth.y", coly = "bot_depth.x")
```

## Compare 2019 to 2021 data (without 2021 observations in 2021 data)

```{r}
# Compare y to x
aa(dat = dat, colx = "surf_temp.x", coly = "surf_temp.y")
aa(dat = dat, colx = "wtcpue.x", coly = "wtcpue.y")
aa(dat = dat, colx = "scientific.x", coly = "scientific.y")
aa(dat = dat, colx = "numcpue.x", coly = "numcpue.y")
aa(dat = dat, colx = "longitude.x", coly = "longitude.y")
aa(dat = dat, colx = "latitude.x", coly = "latitude.y")
aa(dat = dat, colx = "datetime.x", coly = "datetime.y")
# aa(dat = dat, colx = "file.x", coly = "file.y")
aa(dat = dat, colx = "bot_temp.x", coly = "bot_temp.y")
aa(dat = dat, colx = "bot_depth.x", coly = "bot_depth.y")
```

## Compare the datasets in whole

```{r}
comparedf(d2019, d2021)
```

## Compare the datasets without 2021 observations in 2021 data

```{r}
comparedf(d2019, d2021[d2021$year != 2021, ])
```

```{r}
dat_diff <- summary(comparedf(d2019, d2021[d2021$year != 2021, ]))
dat_diff$diffs.byvar.table
```

# Look through 2021 data and see where things are funny...

```{r}
dat <- d2021
# cols <- c("surf_temp", "wtcpue", "scientific", "numcpue", "longitude", "latitude", "datetime", "bot_temp", "bot_depth")
```

## -9999

```{r}
# for (i in cols){
#   wheres_trouble(dat = dat, col = i, trouble = -9999)
# }
trouble <- -9999
wheres_trouble(dat = dat, col = "datetime", trouble = trouble)
wheres_trouble(dat = dat, col = "wtcpue", trouble = trouble)
wheres_trouble(dat = dat, col = "scientific", trouble = trouble)
wheres_trouble(dat = dat, col = "numcpue", trouble = trouble)
wheres_trouble(dat = dat, col = "longitude", trouble = trouble)
wheres_trouble(dat = dat, col = "latitude", trouble = trouble)
wheres_trouble(dat = dat, col = "wtcpue", trouble = trouble)
wheres_trouble(dat = dat, col = "bot_temp", trouble = trouble)
wheres_trouble(dat = dat, col = "bot_depth", trouble = trouble)
```

## NA

```{r}
# for (i in cols){
#   wheres_trouble(dat = dat, col = i, trouble = NA)
# }
trouble <- NA
wheres_trouble(dat = dat, col = "datetime", trouble = trouble)
wheres_trouble(dat = dat, col = "wtcpue", trouble = trouble)
wheres_trouble(dat = dat, col = "scientific", trouble = trouble)
wheres_trouble(dat = dat, col = "numcpue", trouble = trouble)
wheres_trouble(dat = dat, col = "longitude", trouble = trouble)
wheres_trouble(dat = dat, col = "latitude", trouble = trouble)
wheres_trouble(dat = dat, col = "wtcpue", trouble = trouble)
wheres_trouble(dat = dat, col = "bot_temp", trouble = trouble)
wheres_trouble(dat = dat, col = "bot_depth", trouble = trouble)
```

## What stations for each year and survey look like

```{r, echo=FALSE}
dat <- d2021
plot_stations(dat = dat, srvy0 = "EBS")
plot_stations(dat = dat, srvy0 = "NBS")
plot_stations(dat = dat, srvy0 = "BSSLOPE")
plot_stations(dat = dat, srvy0 = "GOA")
plot_stations(dat = dat, srvy0 = "AI")
```


