---
title: 'Access public data using the API and R'
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  md_document: 
    variant: gfm
csl: "https://raw.githubusercontent.com/citation-style-language/styles/master/apa-no-ampersand.csl"
bibliography: "https://raw.githubusercontent.com/afsc-gap-products/citations/main/cite/bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```

### Connect to the API with R

More information about how to amend API links can be found [here](https://docs.oracle.com/en/database/oracle/oracle-rest-data-services/22.3/books.html#AELIG90103/).

Load the first 25 rows of data

```{r api-link, echo = TRUE}
 # install.packages(c("httr", "jsonlite"))
library("httr")
library("jsonlite")
 # link to the API
api_link <- "https://apps-st.fisheries.noaa.gov/ods/foss/afsc_groundfish_survey/"
```

```{r api-all-data-ex, echo = TRUE}
res <- httr::GET(url = api_link)
 # res # Test connection
data <- jsonlite::fromJSON(base::rawToChar(res$content))
 # names(data)
knitr::kable(head(data$items, 3)) 
```

Load the first 10000 rows of data

```{r test-1}
res <- httr::GET(url = paste0(api_link, "?offset=0&limit=10000"))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
dim(data$items)
```

```{r test-2, eval = FALSE, echo = FALSE}
 # Filter by Year: Show all the data greater than the year 2020,
 # https://apps-st.fisheries.noaa.gov/ods/foss/trade_data/?q={"year":{"$gt": 2010}}
res <- httr::GET(url = paste0(api_link, '?q={"year":{"$gt": 2020}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
knitr::kable(summary(data$items[, c("year", "srvy")], 3))
```

Filter by Species name: Show all the data where the product name contains pollock Please note that here the word pollock is case sensitive.

The notation for finding a string is to use % around it. Since % is a reserved character in a URL, you have to replace `%` with `%25`. 

```{r test-3}
res <- httr::GET(url = paste0(api_link, '?q={"common_name":{"$like":"%25pollock%25"}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
knitr::kable(head(data$items, 3))
```


```{r test-4, eval = FALSE, echo = FALSE}
 # Combination of year and name filters: Show all the data where years > 2020 and the product name contains pollock
res <- httr::GET(url = paste0(api_link, '?q={"year":{"$gt": 2020},"name":{"$like":"%25pollock%25"}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
knitr::kable(head(data$items, 3))
```

```{r test-5, eval = FALSE, echo = FALSE}
 # Combination of year, srvy, stratum: Show all the data where year = 1989, srvy = "EBS", and stratum is not equal to 81
 # https://apps-st.fisheries.noaa.gov/ods/foss/trade_data/?q={"year":"1989","srvy": "EBS","stratum":{"$ne": "81"}}

res <- httr::GET(url = paste0(api_link, '?q={"year": 1989,"srvy": "EBS","stratum":{"$ne": "81"}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
knitr::kable(head(data$items, 3))
```

### Subset data

Here, as an example, we can subset the data for the 2018 Aleutian Islands Bottom Trawl Survey. 

```{r api-subset-ex, echo = TRUE, eval = FALSE}
res <- httr::GET(url = api_link, query = list(year = "2018", srvy = "AI"))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
x <- data$items
x <- x[,c("srvy", "year", "stratum", "station", "vessel_name", "latitude_dd", "longitude_dd",
          "species_code", "common_name", "scientific_name", "taxon_confidence",
          "cpue_kgha", "cpue_noha", "weight_kg", "count",
          "bottom_temperature_c", "surface_temperature_c", "depth_m")]
knitr::kable(head(x, 3))
```

```{r child-footer, child=here::here("docs","footer.Rmd")}
```
